### 2024-02-13 doc/15yrsChange.Rmd
Add Study Site, Census and Mortality Descriptions

## Study site

@@

The study was conducted at The Smithsonian Conservation Biology Institute’s (SCBI) large forest dynamics plot in Front Royal, Virginia, USA (38.8935°, -78.1454°; Bourg et al. 2013, Anderson-Teixeira et al. 2015). *A sentence about the region and that it is adjacent to the northern entrance of Shenandoah National Park.* The plot is 25.6 ha (640 x 400 m) and includes a 4-ha, 2.4 m high deer exclusion fence, which has decreased the number/decreased the presence of white-tailed deer since 1990 (McShea 2000). It consists of a mature secondary eastern deciduous forest that developed following logging and agricultural use in the mid-19th century. *Canopy trees are 84-124 years old (Thompson and McGarvey - unpublished data referenced in McGarvey 2013 - see if you can find the published version of this data)* The species composition reflects what is typical of this forest type, with a canopy that is dominated by tulip poplar (*Liriodendron tulipifera*), oak (*Quercus spp.*), and hickory (*Carya spp.*), and the understory that is primarily composed of spicebush (*Lindera benzoin*), paw-paw (*Asimina triloba*), American hornbeam (*Carpinus caroliniana*), and witch hazel (*Hamamelis virginiana*)

*Which definition of forest type is more accurate - mature, mixed deciduous forest or mature secondary eastern deciduous forest?*

@@

@@

*(describe main census & years occurred)*

Data was collected according to the standard ForestGEO protocol, outlined in (Condit 1998). A census crew measured the diameter at breast height (DBH), mapped, and tagged every stem on the plot with a DBH greater than or equal to 1 cm. DBH was recorded only for living trees. If there were multiple stems on a tree, each stem was measured and given a stem number and tag.

*(describe mortality census)*

In addition to the ForestGEO census, an annual [official name of mortality survey] is conducted on the large forest dynamics plot. The survey assesses the health of individual trees and identifies factors that contribute to mortality. The survey was initiated in 2008, and has been conducted every year since 2013. (source). The field crew measured the percentage of the crown that was alive and intact, and noted any factors associated with death (i.e. wounds) present on the tree.  Numerous nuisance species and pathogens, including the emerald ash borer, white-tailed deer, and dutch elm disease have been recorded in the region prior to the start of the mortality survey (Herms & McCullough, 2014, McShea 2000,NEED ELM SOURCE).

*(describe invasive plants survey)*

@@

Field data was collected on Apple iPads using the ESRI ArcGIS Field Maps application. To improve the quality of the data, GitHub Actions continuous integration was implemented for quality control (Kim et al., 2022). Data from the previous census was used to populate the Field Maps app, and fields were created to input data. There were checks built into the app to help catch errors, for example the DBH field would flag a suspiciously high number and would not allow the form to be submitted if it was blank. To use the app in the field the census crew downloaded offline areas of where they were working and synced them to an online map after collecting data. Next the data was downloaded from ArcGIS online and pushed to GitHub, where it triggered a GitHub Action running data quality assurance and control (QA/QC) scripts. The scripts were written in R and checked the data for errors such as “suspicious positive growth.” After going through the QA/QC process, a csv of errors and warnings with the associated stem were output which could then be corrected in the field.

>>>>

## Site Description

@@

@@

The study was conducted in the 25.6-hectare Large Forest Dynamics plot at the National Zoo and Conservation Biology Institute in Front Royal, Virginia (38 53’36.6“N, 78 08’43.4 “W).  The plot, which is located in the central Appalachian Mountains adjacent to Shenandoah National Park, is composed of secondary eastern mixed deciduous forest. Situated in the Appalachian Oak forest region, the canopy is dominated by tulip poplar (*Liriodendron tulipifera*), oak (*Quercus spp.*), and hickory (*Carya spp.*) and the understory is primarily composed of spicebush (*Lindera benzoin*), paw-paw (*Asimina triloba*), American hornbeam (*Carpinus caroliniana*) and witch hazel (*Hamamelis virginiana*) (SCBI site description). The land-use history of the site is varied, including periods of agricultural development and intensive logging, with dendrological data estimating canopy tree establishment around 1900 (SI archives, Bourg et al. 2013).  Natural disturbances at the plot consist of wind and infrequent ice storms (Anderson-Teixiera 2015).  The plot, which includes a 4-hectare deer exclosure that has decreased the presence of deer since 1990, is divided into 640 quadrats, each measuring 20 x 20 meters. It is one of 78 sites in the Forest Global Earth Observatory (ForestGEO), a global network of forest dynamic plots that promotes comparative forest ecology studies around the World.

@@

*(Describe main census & years occurred)*

A number of surveys are regularly conducted at the site, four of which are included in our analysis.  As part of the ForestGEO network, the plot undergoes a comprehensive woody plant inventory every five years, according to the protocol outlined in Condit (1998).  This inventory, hereby referred to as the census, includes all stems greater than 1 centimeter in diameter at 1.3 meters in height (referred to as diameter at breast height or dbh).  The census records information regarding the dbh, species, status and location of each stem included in the survey.  Each plant is assigned an identifying number for sequential data collection in subsequent censuses and outfitted with a metal tag in the field. For multi-stemmed individuals, each additional stem that surpasses the 1 cm dbh threshold receives a stem number and associated tag.  The location of each individual within its respective 20 x 20 meter quadrat is recorded on a map of the plot on ArcGIS FieldMaps.  Established in 2008, there have been 4 censuses at the site, comprising 20 years of detailed forest dynamics data.

*Is dbh considered common knowledge? Does it need to be defined?*

*Check with Luca to use consistent terminology here and in the analysis portion of the methods for ID tag, stem tag, etc.*

*(Describe mortality census and invasive plant survey)*

In addition to the ForestGEO census, an annual mortality survey is conducted on the site.   Through assessing individual tree health, this study illustrates trends in forest mortality and identifies factors that are associated with death.  Data is collected on the current status, canopy position, percentage of crown living and intact, and visible indicators of poor tree health, such as physical damage, potential pathogens and insect infestation.  Furthermore, an invasive plant survey is conducted in conjunction with [the mortality survey or the census].  This study evaluates the establishment of non-endemic plant species and patterns of spatial distribution. Quadrats are visually examined for non-endemic plant species, with an estimation of the area covered by each species represented on a 1-5 numeric scale.

*Is the invasive plant survey conducted every year or every five years?*

@@

To improve the quality of the data, GitHub Actions continuous integration was implemented for quality control (Kim et al., 2022). Data from the previous census was used to populate the Field Maps app, and fields were created to input data. There were checks built into the app to help catch errors, for example the DBH field would flag a suspiciously high number and would not allow the form to be submitted if it was blank. To use the app in the field the census crew downloaded offline areas of where they were working and synced them to an online map after collecting data. Next the data was downloaded from ArcGIS online and pushed to GitHub, where it triggered a GitHub Action running data quality assurance and control (QA/QC) scripts. The scripts were written in R and checked the data for errors such as “suspicious positive growth.” After going through the QA/QC process, a csv of errors and warnings with the associated stem were output which could then be corrected in the field.
